---
title: "Introduction to spectral processing with 'spectrolab'"
author: "Anna Schweiger"
output:
  # pdf_document: default
  html_document: default
---

## Install spectrolab

```{r, eval=FALSE}
install.packages("spectrolab")
```

```{r, message=F}
library("spectrolab")
```

## The spectra class

`spectrolab` defines a new S3 class called `spectra` that holds all of the different components of a spectral data.

Without diving too much into its implementation, a `spectra` object holds the important information needed for most spectral data sets: reflectance, wavelengths, filen ames, metadata etc. The class has a bunch of requirements in terms of both format and values. 

## Read and inspect data
Our spectral data were measured with an instrument called ASD.
```{r, message=F}
spec <- read_spectra("./example_data/", format="asd")

```

`spectrolab` can also read other file formats, but let's not worry about that for now. You can always look at `spectrolab's` help for more information. 

```{r, eval=F}
help(read_spectra)
```

You can see what a `spectra` object contains by typing

```{r}
spec
```
Our `spectra` object contains 9 samples, > 2000 bands, and no metadata. To access the individual components try

```{r,eval=F}
bands(spec)
tail(bands(spec))

names(spec)
meta(spec)
```

Let's plot our spectra. It's as simple as

```{r}
plot(spec)
```

You can add spectral regions and quantiles to the plot. How? Try to figure it out yourself. Hint: you can search the help page, or type _**comma**_ after _**spectrolab**_ followed by the  _**TAB**_ key

```{r}
help(package="spectrolab")
```

Try also `spectrolab's` interactive plotting option

```{r,eval=FALSE}
plot_interactive(spec)

```
If you'd like to remove a spectrum for some reason, click on it and note the number.  Let's create a new `spectra` object and remove the spectrum with the highest reflectance

```{r}
spec2 <- spec[-4]

# standard plotting commands can be added
plot(spec2, col=rainbow(n=length(spec2)), lwd=2)
```

## Adding metadata 

Easy! How about adding some metadata?  One option is to edit the `spectra` object.
```{r}
meta(spec2, label = "some_cool_data") <- c(3,7,4,2,6,9,8,7)
```
Adding a dummy N content
```{r}
n_content <- rnorm(n = nrow(spec2), mean = 2, sd = 0.5)
meta(spec2, label = "N_percent") = n_content
# and check 
meta(spec2)
```
Looks good! You can also import a metadata table and merge it to your spectral data. 

```{r}
meta_csv <- read.csv("./example_data/metadata_spec2.csv")
meta(spec2) <- meta_csv[,c(2,3)]

meta(spec2)
```
Awesome! You can query your `spectra` object and metadata by name or index
```{r}
meta(spec2, sample=c(1,3))

names(spec2)
meta(spec2,sample="103_SALHU00000")
```
Bonus question: What's the difference in the output of these two functions
```{r,eval=FALSE}
meta(spec2,label = 2)
meta(spec2,label = 2, simplify = TRUE)
```

## Basic calculations
Sometimes you might want to calculate a mean spectrum, or other things

```{r}
spec_mean <- mean(spec2)

plot(spec2)
plot(spec_mean, col="red", lwd=3, add=T)
plot_quantile(spec2,total_prob = 0.75,add=T) 
```
Conveniently, this calculates also the mean of your metadata
```{r}
meta(spec_mean)
```
What about minimum and maximum?
```{r,eval=FALSE}
spec_min <- min(spec2) 
meta(spec_min)
```
This does not work because `spectrolab` does not know if you want a minimum spectrum, a minimum reflectance value, a minimum values of a specific metadata element, etc.  But there are better ways to calculate each of those... can you figure them out?

## Some post-processing options
You might want to cut off noisy regions at the beginning and end of the spectrum. Simple sub-setting does the trick. Use `plot_interactive` to decide which regions are "good"

```{r}
spec_trim <- spec2[ , bands(spec2, 500, 2400)]
```
You can also re-sample wavelengths, e.g. to reduce the amount of data
```{r}
spec_sub <- resample(spec2, new_bands = seq(400, 2400, 10))
```
Or select specific samples
```{r}
spec3 <- spec2[meta(spec2,"N_percent")>2,]
meta(spec3, "N_percent")

spec4 <- spec2[grepl("SPITO",names(spec2))]
spec4
```

## Export data
You can of course export and re-import your data. Just be careful how `spectrolab` imports tabular data. You will need to specify names, bands and metadata

```{r}
write.csv(spec_sub, "./example_data/export_spectra.csv", row.names = F)

spec_csv <- read.csv("./example_data/export_spectra.csv")
```
`spectrolab's` bands names need to be numeric. So we need to modify our column names first

```{r}
names(spec_csv) <- gsub("X","",names(spec_csv))
spec5 <- as_spectra(spec_csv,name_idx = 1, meta_idxs = 2:5)

plot(spec5, col=rainbow(n=length(spec5)), lwd=6)
```
**Beautiful!!**

### Wishing you lots of fun with all your spectral endevours! 

